---
title: "SRP_Assessment2"
author: "Sandeep Kumar Patel (S4651719)"
date: '2022-03-27'
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 2: Data Wrangling and visualization

```{r}
#install.packages("tidyverse")
library(tidyverse)

#install.packages("knitr")
library(knitr)

#install.packages("skimr")
library(skimr)

#install.packages("dplyr")
library(dplyr)

thankgiving_meals <-read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-20/thanksgiving_meals.csv")

```


### 1. Display the first 10 rows of the dataset using `kable()` function (1 marks).

```{r}
knitr::kable(head(thankgiving_meals[1:10,]), "pipe")
```

### 2. Using `skim()` display the summary of variables. 

#### Summary of variables Information

```{r}
skim(thankgiving_meals)
```


### Think about the task to predict a family income based on their menu: what variables may be useful? Are all of them correct type? 

### Write 2-3 sentences with your explanation. (2 marks)

The Variables that can be useful to predict the family income of the household are main dish, main prep, stuffing, carnebrrey, and gravy. This can justified as the missing values are 84 which is equal to the individual those who do not celebrate thanksgiving.

### Think about the task to predict a community type or US_region based on their menu: what variables may be useful? Are all of them correct type? (2 marks)

Community type or US region can be predicted by variables such as side dishes, pie, dessert,  and income. 


### 3. Use `fct_reorder` and `parse_number` functions to create a factor variable `family_income`(2 mark).

```{r}
family_income <- thankgiving_meals %>%
  mutate(family_income = fct_reorder(family_income, parse_number(family_income))) %>%
  filter(!is.na(family_income))
```

### 4. What is the number of people who celebrate? (1 mark)

```{r}
thankgiving_meals %>%
  filter(celebrate=="Yes")%>%
  count(celebrate)
```

From the total sample size 980 individuals celebrate thanksgiving.

### 5. What are categories and insights for each main dish served and the method it is prepared? (2 marks)

```{r}
kable(thankgiving_meals %>%
        count(main_dish, main_prep, sort = TRUE) %>%
        filter(!is.na(main_dish) & !is.na(main_prep)), format = "html")
```

### 6. Create 3 different data viz showing insights for main dish served and the method. Provide your own legend and use themes.
### Write 2-3 sentences with your explanation of each insight. (4 marks)

#### First Data visual Presentation

```{r}
thankgiving_meals %>%
  filter(main_prep == "Baked" | main_prep == "Fried" | main_prep == "Roasted") %>%
  ggplot(aes(main_dish, fill = main_dish)) + geom_bar() + facet_wrap(vars(main_prep)) +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

#### Explanation

This Visualization helps use to understand how the main dish is prepared. This also concluded what kind of meat is prepared. roasted or baked turkey is the most common main dish.

#### Second Data visual Presentation

```{r}
thankgiving_meals%>%
  ggplot(aes(  main_dish, main_prep))+
  geom_jitter(height=2, width=2.5,color="blue",shape=0)+
  scale_y_discrete()+
  labs(title="Second Visualization")+
  coord_flip()
```

#### Explanation

The Dots in the graph represents the individual observations and how they like to prepare and what meat they like. This can help to grasp the individual observations.

#### Third Data visual Presentation

```{r}
thankgiving_meals%>%
  ggplot(aes(  main_dish, main_prep))+
  geom_count(color="purple",size=1)+
  labs(title="Third Visualization")+
  coord_flip()
```

#### Explanation

This graph gives a brief overview on the link between prep method and meat used. It can seen that no one prefers to fry tofurkey or bake turducken.

### 7. How many use cranberry sauce? How many use gravy? 2marks

##### Cranberry

```{r}
kable(thankgiving_meals %>%
        filter(!is.na(cranberry) & cranberry!="None") %>%
        count(), "html")
```

There are 828 individuals that like to use carnberry sauce.

##### Gravy

```{r}
kable(thankgiving_meals %>%
        filter(gravy=="Yes") %>%
        count(gravy), "html")
```

There are 892 individuals that like to use gravy.

### 8-9. What is the distribution of those who celebrate across income ranges. Create a data viz.
### Write 2-3 sentences with your explanation of each insight. (4 marks)

```{r}
table(thankgiving_meals$family_income)
thankgiving_meals %>%
  filter(celebrate == "Yes") %>%
  ggplot(aes(family_income, fill=family_income)) + geom_bar() + facet_wrap(vars(family_income)) + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  ggtitle("Thanksgiving Celebrations")
```

##### Explanation

Frome the Visual representation it can be concluded that family income from 25,000 to 125,000 are the individuals that celebrate thanksgiving more than other income range.

### 10. Use the following code to create a new data set 2 mark

`
select(id, starts_with("side"),
         starts_with("pie"),
         starts_with("dessert")) %>%
  select(-side15, -pie13, -dessert12) %>%
  gather(type, value, -id) %>%
  filter(!is.na(value),
         !value %in% c("None", "Other (please specify)")) %>%
  mutate(type = str_remove(type, "\\d+"))
`

### Write 2-3 sentences with your explanation of what it does. (4 marks)

```{r}
thanksgiving_new <- thankgiving_meals%>%
  select(id, starts_with("side"),
         starts_with("pie"),
         starts_with("dessert")) %>%
  select(-side15, -pie13, -dessert12) %>%
  gather(type, value, -id) %>%
  filter(!is.na(value),!value %in% c("None", "Other (please specify)")) %>%
  mutate(type = str_remove(type, "\\d+"))
```

##### Explanation
The above codes extract three variables Side dish, Pie, and Dessert from the data ommitting the individual the that have not aswered other (please specify) variable. and extracts Information such as ID, Value, and Type.

### 11-12. Install package `widyr` and use `pairwise_cor()` function https://www.rdocumentation.org/packages/widyr/versions/0.1.3/topics/pairwise_cor
### Write 2-3 sentences with your explanation of what it does. (2 marks)

### Use this code for the new dataset

`
pairwise_cor(value, id, sort = TRUE)

`
### Write 1 sentence with your explanation of what insights it shows. (2 marks)

```{r}
#install.packages("widyr")
library(widyr)

thanksgiving_new%>%
  pairwise_cor(value, id, sort = TRUE)
```

##### Explanation
Thiese codes provides that correlation coefficent of the variable that were extracted from datasets.

### 13. Use `lm()` or randomForest() function to build a model that predict a family income based on data in the dataset. 8 marks

### Compare 3 models using different set of input variables. Use different number of variables.

### Explain your choice of variables (3 sentences) 

### Write 2 sentences explaining which model is best.




