---
title: "SRP Assessment 1"
author: "Sandeep Kumar Patel s4651719"
date: '2022-03-06'
output: html_document
---

*2 Create the heading of the first level and call it RMarkdown editing*

# Rmarkdown editing

*3. Write 2-3 sentences about RMarkdown (you can google this information or use resources recommended in class - no need to reference)*

*4. In the above use bold and italics for editing.*

**R Markdown** provides an authoring framework for data science. You can use a single R Markdown file to both

*1. Save and execute code*
*2. Generate high quality reports that can be shared with an audience*

R Markdown documents are fully reproducible and support dozens of static and dynamic output formats.

*5. Review the suggested documentation on how to insert links in the .rmd file and include an in-text link to https://rmarkdown.rstudio.com/lesson-8.html*

[Click here](https://rmarkdown.rstudio.com/lesson-8.html)

*6. Insert an R chunk and create a variable with this dataset*
*https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv*

*The name of the variable should be `fastfood`*

```{r}
#installpackages(tidyverse)
library(tidyverse)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

```

*7. Display the first 10 rows of the dataset using head() and kable().*

*Display the first 10 rows of the dataset and the first 5 variables*

*Use Help and the link below to read more about those functions  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html*

## Use of Head function

```{r}
fastfood[1:10, 1:5]
```


## Use of kable function

```{r}
knitr::kable(head(fastfood[1:10, 1:5]), "pipe")
```

*8. Display the observations that has more than 1000 calories*

```{r}
filter (fastfood, calories>1000)
```

*9. Arrange observations with more than 40 in total_fat and more than 80 in total_carb in the descending order (PLEASE USE THE VARIABLE OF YOUR CHOICE TO ORGANISE THE DESCENDING ORDER) and save them to a new variable (dataset) called `dont_eat_this`*

*Display & total fat in descending order *

```{r}
fastfood %>% filter (total_fat> 40 & total_carb> 80) %>% 
  arrange(desc(total_fat))
```

##Save

```{r}
dont_eat_this <-fastfood %>% filter(total_fat> 40 & total_carb> 80)
```

*10. Using the initial dataset variable, use `case_when()` to create a new variable `heavy_food` which is equal to "heavy" when total calories (`calories`) are greater than 500, "low" when total calories are less than 250 and "average" for all other cases.*

```{r}
heavy_food <-  fastfood %>% 
  group_by(calories)%>%
  mutate (heavy_food = case_when(calories >500 ~ "Heavy", calories < 250  ~ "Low", TRUE ~ "Average")) %>% 
  ungroup()
```

*Count the number of observations in each created category.*

```{r}
heavy_food %>%
  count(heavy_food, wt = NULL, sort = TRUE, name = "observations")
```

*11. Display the types of variables in the dataset using `skimr` package*

```{r}
#installpackages(skimr)
library(skimr)
skim(fastfood)
```

*12. Present the count observations from each restaurant in a descending order*

```{r}
fastfood %>%
  count (restaurant, wt = NULL, sort = TRUE, name = "res_num")%>%
  arrange(desc(restaurant))
```

*13. Using groupings (group_by()), summarise and display the average number of calories for each restaurant.*

```{r}
fastfood%>%
  group_by(restaurant)%>%
  summarise(average_cal = mean(calories, trim = 0, na.rm = FALSE)) %>% 
  ungroup() 
```

*14. Add variables to the dataset, which:*

##Calculates the average calories per type of restaurant and call it `average_calories`*

```{r}
average_calories <-  fastfood %>% 
  group_by(restaurant)%>%
  mutate(average_calories = mean(calories)) %>% 
  ungroup()
```

##Calculates the maximum total_fat per type of restaurant and call it `max_fat`*

```{r}
max_fat <-  fastfood %>% 
  group_by(restaurant)%>%
  mutate(max_fat = max(calories)) %>% 
  ungroup()
```


##Calculates the minimum cholesterol per type of restaurant and call it `min_cholesterol`

```{r}
min_cholesterol <-  fastfood %>% 
  group_by(restaurant)%>%
  mutate(min_cholesterol = min(calories)) %>% 
  ungroup()
```

*15. Display the data vis of total fat per each type of restaurant. Write a narration (2-3 sentences) why you believe this type of data viz presents such information best.*

##Display

```{r}
fastfood %>% 
  group_by(restaurant) %>% 
  ungroup() %>% 
  ggplot(aes(x = restaurant, y = total_fat)) +
  geom_col() +
  labs(x="Names of Restaurant", y="Total Fat", title="Data Visualisation 1", caption = "Image 1:Bar Graph")
```

##**Write**

Before we being to type the codes it is important to identify which variable is what either it is discrete or contiunous or both respectively. By using to the cheatsheet we can identify which can be used to best of its ability.*

*16. Add a variable to the dataset, which calculates the sum of cholesterol and sodium and call it `cholesterol_sodium`.*

```{r}
cholesterol_sodium <- fastfood %>% 
  group_by(restaurant) %>%
  mutate(sum_cholesterol_sodium = cholesterol + sodium) %>% 
  ungroup()
```

```{r}
library(dplyr)

cholesterol_sodium %>%
  select(-salad)
```

*17. Use observations for Mcdonalds to plot sugar variable against protein with `geom_point()'*

```{r}
fastfood %>% 
  select(restaurant, sugar, protein) %>% 
  filter(restaurant=="Mcdonalds") %>% 
  ggplot(aes(sugar,protein)) +
  geom_point(colour = "red") +
  labs(x="Sugar", y="Protein", title = "Mcdonalds - Visualisation", caption = "Point Graph")
```

*Part 3. - Due session 8 - continue in the same file, but please include "Part 3" as a heading*

**Focus**: *ggplot2, factors, strings, dates*

*18. Identify variable(s) which should be factors and transform their type into a factor variable.*

```{r}
#install.packages("forcats")
library(forcats)
#is.factor(fastfood$restaurant)
#as.factor(fastfood$restaurant)
```

*19. Create a new variable:*

*Read about `cut_number()` function using Help and add a new variable to the dataset `calories_type`. Use `calories` variable for `cut_number()` function to split it into 3 categories `n=3`, add labels `labels=c("low", "med", "high")` and make the dataset ordered by arranging it according to calories.*

```{r}
calories_type <- fastfood %>% 
  group_by(restaurant) %>% 
  mutate(calories_type = cut_number(calories, n = 3, labels = c("low","med","high"))) %>% 
  ungroup() %>% 
  arrange(calories_type)
```

*20. Create a dataviz that shows the distribution of `calories_type` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful. Use factors related data viz functions.*

```{r}
calories_type %>% 
  group_by(restaurant) %>% 
  ungroup() %>% 
  ggplot(aes(restaurant, calories_type)) +
  geom_boxplot(colour="red") +
  facet_wrap(vars(restaurant)) +
  coord_flip() +
  theme_bw() +
  labs(x="Restaurant", y="Calories Type", title ="Data Visualisation", caption = "Plot Graph")
```

*21.  Add a new variable that shows the percentage of `trans_fat` in `total_fat` (`trans_fat`/`total_fat`).  The variable should be named `trans_fat_percent`. Do not forget to save the updated dataset.*

```{r}
fastfood <- fastfood %>% 
  group_by(restaurant) %>% 
  mutate(trans_fat_percent = trans_fat / total_fat * 100) %>% 
  ungroup()
```

*22. Create a dataviz that shows the distribution of `trans_fat` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful.*

```{r}
ggplot(fastfood, aes(restaurant,trans_fat)) +
  geom_point(colour = "red") +
  facet_wrap(vars(restaurant)) +
  coord_flip() +
  labs(x="Restaurant", y="Trans Fat", title = "Data Visualisation", caption = "Point Graph") +
  theme()
```

*23. Calculate and show the average (mean) `total_fat` for each type of restaurant. No need to save it as a variable.*

```{r}
fastfood %>% 
  group_by(restaurant) %>% 
  summarise(average_fat = mean(total_fat)) %>% 
  ungroup()
```

*24. And create a dataviz that allow to compare different restaurants on this variable (`total_fat`). You can present it on one dataviz (= no facets).*

```{r}
fastfood %>% 
  group_by(restaurant) %>% 
  summarise(average_fat = mean(total_fat)) %>% 
  ungroup() %>% 
  mutate(average_fat = fct_reorder(restaurant,average_fat)) %>% 
  ggplot(aes(restaurant,average_fat)) +
  geom_col()
```












