---
title: "SRP Assessment 1"
author: "Sandeep Kumar Patel s4651719"
date: '2022-03-06'
output: html_document
---

*2 Create the heading of the first level and call it RMarkdown editing*

# Rmarkdown editing

*3. Write 2-3 sentences about RMarkdown (you can google this information or use resources recommended in class - no need to reference)*

*4. In the above use bold and italics for editing.*

**R Markdown** provides an authoring framework for data science. You can use a single R Markdown file to both

*1. Save and execute code*
*2. Generate high quality reports that can be shared with an audience*

R Markdown documents are fully reproducible and support dozens of static and dynamic output formats.

*5. Review the suggested documentation on how to insert links in the .rmd file and include an in-text link to https://rmarkdown.rstudio.com/lesson-8.html*

[Click here](https://rmarkdown.rstudio.com/lesson-8.html)

*6. Insert an R chunk and create a variable with this dataset*
*https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv*

*The name of the variable should be `fastfood`*

```{r}
#installpackages(tidyverse)
library(tidyverse)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')

```

*7. Display the first 10 rows of the dataset using head() and kable().*

*Display the first 10 rows of the dataset and the first 5 variables*

*Use Help and the link below to read more about those functions  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html*

## Use of Head function

```{r}
fastfood[1:10, 1:5]
```


## Use of kable function

```{r}
knitr::kable(head(fastfood[1:10, 1:5]), "pipe")
```







